 Casting multiple shadows: interactive data visualisation with tours and embeddings

### 2020-02-06 paper.Rmd
rejigging and write up about methods used.

@@

Non-linear methods
generally perform preprocessing on the high-dimensional data such as generating
a neighborhood graph and perform transformations on the preprocessed form. We
restrict our attention to three methods that are commonly used in
high-throughput biology: t-distributed stochastic neighbor embedding (t-SNE), 
uniform manifold alignment and projection (UMAP), and PHATE. The t-SNE algorithm estimates the similarity of (Euclidean) distances of points in a high dimensional space using a Gaussian distribution and then  
estimates a configuration in the low dimensional embedding space by modelling
similarities using a t-distribution with 1 degree of freedom. The resulting
configuration is the one that minimizes the Kulback-Leibler divergence
between the two distributions. A recent theoretical contribution by @Linderman2019-dq proved that t-SNE can recover spherical and well separated 
cluster shapes, and proposed new approaches for tuning the optimisation parameters. It is a known problem that t-SNE can have trouble recovering
global structure and that configurations can be highly dependent on
how the algorithm is initialised and parameterised 
[@Wattenberg2016-ji; @Kobak2019-lm]. UMAP is a method that is
related to LargeVis, and like t-SNE acts on the k-nearest neighbor graph. Its
main differences are that it uses a different cost function (cross entropy) which is optimized using stochastic gradient descent and defines a different
kernel for similiarites in the low dimensional space. Finally, PHATE is based

>>>>

Non-linear methods
generally perform preprocessing on the high-dimensional data such as generating
a neighborhood graph and perform transformations on the preprocessed form. We
restrict our attention to three methods that are commonly used in
high-throughput biology: t-distributed stochastic neighbor embedding (t-SNE), 
uniform manifold alignment and projection (UMAP), and potential of heat-diffusion for affinity-based transition embedding (PHATE). The t-SNE algorithm estimates the similarity of (Euclidean) distances of points in a high dimensional space using a Gaussian distribution and then  
estimates a configuration in the low dimensional embedding space by modelling
similarities using a t-distribution with 1 degree of freedom. The resulting
configuration is the one that minimizes the Kulback-Leibler divergence
between the two distributions. A recent theoretical contribution by @Linderman2019-dq proved that t-SNE can recover spherical and well separated 
cluster shapes, and proposed new approaches for tuning the optimisation parameters. It is a known problem that t-SNE can have trouble recovering
global structure and that configurations can be highly dependent on
how the algorithm is initialised and parameterised 
[@Wattenberg2016-ji; @Kobak2019-lm]. UMAP is a method that is
related to LargeVis, and like t-SNE acts on the k-nearest neighbor graph. Its
main differences are that it uses a different cost function (cross entropy) which is optimized using stochastic gradient descent and defines a different
kernel for similiarites in the low dimensional space. Due to it's computational
speed it's possible to generate UMAP embeddings in more than three dimensions. Finally, PHATE, inspired by diffusion maps, is based on estimating an affinity matrix via a distance matrix and k-nearest neighbors graph. The algorithm
denoises estimated distances in high dimensional space via transforming the
affinity matrix into a Markov transition probaility matrix and diffusing this 
matrix over a fixed number of time steps. Then the diffused probabilities
are transformed once more to construct a distance matrix, and multidimensional
scaling is performed to construct a 2d embedding for visualization.

@@
